---
title: "R Notebook"
output: html_notebook
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
library(dplyr)
library(tidytext)
library(ggplot2)
library(stringr)
library(gutenbergr)
library(tidyverse)
library(scales)
library(gmodels)
```

# H.G. Wells and Bront&euml; sisters
## Data Wrangling and Statistical Analysis

H. G. Wells books: The Time Machine, The War of the Worlds, The Invisible Man, and The Island of Doctor Moreau. 

```{r}
# Download HG Wells books from Gutenberg
hgwells <- gutenberg_download(c(35, 36, 5230, 159))
tidy_hgwells <- hgwells %>%
  # the text of the four HG Wells books are separated into word tokens
  unnest_tokens(word, text) %>% 
  # anti_join with stop_words removes stop words
  anti_join(stop_words)
```

Are any words not lowercase characters that may be in single quotes?

The function `str_detect()` can detect the presence or absence of a pattern in a string.

The pattern `[a-z']+` is another **regular expression**.  What does it mean?

```{r}
sum(str_detect(tidy_hgwells$word, "[a-z']+") == FALSE)
```

What are the most common words in these novels?

```{r}
tidy_hgwells <- tidy_hgwells %>% mutate(word = str_extract(word, "[a-z']+")) %>% 
  count(word,sort = T) 
tidy_hgwells
```

OK, lets do the same thing with all the texts of the Brontë sisters. How could we do this?


```{r}

#Write your code to download Bronte sisters texts here
#Store the Bronte books in a variable called 'bronte' & your tokens in 'tidy_bronte'

bronte <- gutenberg_download(c(1260, 768, 969, 9182, 767))
tidy_bronte <- bronte %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)


```

What are the most common words in these novels of the Bront&euml; sisters?

```{r}
tidy_bronte <- tidy_bronte  %>% mutate(word = str_extract(word, "[a-z']+")) %>% 
  count(word,sort = T) 
tidy_bronte
```

Let's calculate the relative frequencies for each word for the works of Jane Austen and the Bront&euml; sisters and plot.

```{r,cache=TRUE}
comparison <- tidy_bronte %>% inner_join(tidy_hgwells,by="word")  %>% mutate(freq_bronte=n.x/sum(n.x),freq_wells=n.y/sum(n.y))
ggplot(comparison, aes(freq_bronte,freq_wells)) + geom_point() + 
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1, hjust = 1) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

What have we plotted? How can we interpret this plot? Is it meaningful? 

All the words that appear in both texts.  But, among all words in the text how many appear in one text and not the other?

```{r}
comparison_all <-
  tidy_bronte %>% full_join(tidy_hgwells, by = "word") %>%
  mutate(br_in = if_else(is.na(n.x), 0, 1),
  well_in = if_else(is.na(n.y), 0, 1))
  with(comparison_all, CrossTable(br_in, well_in, prop.chisq = F))
```

How do we interpret this table? What is it telling us?


This table tells us that there are 11,710 words in Well's works, and 8,307 (71%) of these words appear in Bront&euml; sisters works.  Conversely, there are 22,678 words that appear in Bront&euml; sisters' works and 8,307 (37%) of these words appear in Well's works.  

- If a word is randomly selected in one of the Well's novels then there is a 0.71 chance that it will also appear in the Bront&euml; sisters works.  This is also called the conditional probability that the word appears in Bront&euml; sisters *given that* it appears in Wells.

- If a word is randomly selected in one of the Bront&euml; sisters works then there is a 0.37 chance that it will also appear in the Well's novels.  This is also called the conditional probability that the word appears in Wells *given that* it appears in Bront&euml; sisters.

Did Wells influence the Bront&euml; sisters?

## Term Frequency

The term frequency in a document is number of times a term $\text t$ occurs in document $\text d$,

$$\text{tf}_\text{t,d}.$$ 

## Inverse Document Frequency

- The inverse document frequency (IDF) is a statistical weight used for measuring the importance of a term in a text document collection. The document frequency DF of a term is defined by the number of documents in which a term appears.

- Karen Sparck-Jones first proposed that terms with low document frequency are more valuable than terms with high document frequency during retrieval (Sparck-Jones K., 1972.). 

- In other words, the underlying idea of IDF is that the more frequently the term appears in the collection, the less informative the term is. Is this intuitively true? Always true? 

The *inverse document frequency* of a term $\text t$ is,

$$\text{idf}_\text{t}=\log\left(\frac{N}{\text{df}_\text{t}}\right).$$ 

$N$ is the total number of documents in a collection (or corpus) of documents, and $\text{df}_\text{t}$ is the number of documents in a collection that contain the term $\text t$.

## N-Grams

A unit of textual analysis, where N is some number. Bigrams: N = 2. Trigrams: N = 3.

# Jane Austen's novels

Let's calculate some of these things by looking at Jane Austen's novels. 


```{r}
library(janeaustenr)
book_words <- austen_books() %>%
  unnest_tokens(word, text) %>%
  count(book, word, sort = TRUE) %>%
  ungroup()
total_words <- book_words %>% 
  group_by(book) %>% 
  summarize(total = sum(n))
book_words <- left_join(book_words, total_words)
book_words
```

Uh-oh!

Let's look at some Jane Austen n-grams:

```{r}

austen_bigrams <- austen_books() %>% unnest_tokens(bigram, text, token = "ngrams", n = 2)
austen_quadgrams <- austen_books() %>% unnest_tokens(quadgram, text, token = "ngrams", n = 4)
austen_quadgrams

austen_bigrams %>% count(bigram, sort = TRUE)
austen_quadgrams %>% count(quadgram, sort = TRUE)

austen_quadgrams
```

Let's count the most common bigrams in Austen's work:

```{r}
bigrams_separated <- austen_bigrams %>% separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>% filter(!word1 %in% stop_words$word) %>% filter(!word2 %in% stop_words$word)

bigram_counts <- bigrams_filtered %>% count(word1, word2, sort = TRUE)

bigram_counts
```

Now let's try mapping the counts of these bigrams in relation to TF-IDF:

```{r}
bigrams_united <- bigrams_filtered %>%
  unite(bigram, word1, word2, sep = " ")

bigram_tf_idf <- bigrams_united %>%
  count(book, bigram) %>%
  bind_tf_idf(bigram, book, n) %>%
  arrange(desc(tf_idf))

bigram_tf_idf
```

## Tf-idf Weighting

A weight for each term in each document is given by multiplying term frequency and inverse document frequency.

$$\text{tf-idf}_\text{t,d}= \text{tf}_\text{t,d} \times \log\left(\frac{N}{\text{df}_\text{t}}\right).$$

Some properties of Tf-idf [(see Manning et al.)](https://nlp.stanford.edu/IR-book/html/htmledition/tf-idf-weighting-1.html):

1. highest when $t$ occurs many times within a small number of documents (thus lending high discriminating power to those documents);
2. lower when the term occurs fewer times in a document, or occurs in many documents (thus offering a less pronounced relevance signal);

3. lowest when the term occurs in virtually all documents.

# Jane Austen's novels

```{r}
book_words <- austen_books() %>%
  unnest_tokens(word, text) %>%
  count(book, word, sort = TRUE) %>%
  ungroup()

total_words <- book_words %>% 
  group_by(book) %>% 
  summarize(total = sum(n))

book_words <- left_join(book_words, total_words)

book_words
```

- One row for each word-book combination.

We can calculate tf-idf for the Jane Austen novels.

```{r}
book_words <- book_words %>% bind_tf_idf(word, book, n)
book_words
```

Let's look at terms with high tf-idf in Jane Austen's works.

```{r}
book_words %>%
  dplyr::select(-total) %>%
  arrange(desc(tf_idf))
```

